---
title: "Colorectal cancer incidence and mortality"
author: "Taavi Päll"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Colorectal cancer incidence and mortality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

## Download list of available datasets
By default, datasets table is loaded from a local copy supplied with the package (`local = TRUE`).
Set argument `local = FALSE` to download fresh list of tables from TAI.
Table columns **Database** and **Node** describe table address in the database tree.
Whereas, database and node names are available only in Estonian.
**Name** column contains table names and should be used when downloading data.
**Title** columns contains table descriptions and in many cases also what are the values presented in the table (e.g. age-normalised incidence per 100'000).


```{r}
library(dplyr)
library(stringr)
library(boulder)
tabs <- get_all_tables(lang = "en")
tabs
```

Find tables with cancer incidence and mortality data. 
Table info is stored in **Title** column.
For better matching we convert Title column to lower case.
Cancer morbidity is stored under 

```{r}
tabs %>% 
  mutate(descr = str_to_lower(Title)) %>% 
  filter(str_detect(descr, "neoplasm")) %>% 
  select(Name, Title, Node, Updated) %>% 
  knitr::kable()
```

We are interested in table **PK30**
```{r}

```

